---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dobEstimator

<!-- badges: start -->
[![R-CMD-check](https://github.com/bairdj/dobEstimator/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/bairdj/dobEstimator/actions/workflows/check-standard.yaml)
[![codecov](https://codecov.io/gh/bairdj/dobEstimator/graph/badge.svg?token=3IZA9U35ZY)](https://codecov.io/gh/bairdj/dobEstimator)
<!-- badges: end -->

The goal of dobEstimator is to allow the estimation of an individual's date of birth, when only
their ages at a set of dates are known. For example, you may have a dataset like this:

```{r test-date, echo = FALSE, warning=FALSE}
library(lubridate, warn.conflicts = FALSE)
age_dataset <- data.frame(
  age = c(20, 20, 21, 23, 23),
  date = as.Date(c("2010-03-05", "2010-05-06", "2010-07-08", "2012-06-23", "2013-01-01"))
)

age_dataset
```

From this data, it can be deduced that the individual's 21st birthday occurred at some
point between 2010-05-07 and 2010-07-08, thus their birth date must be between 1990-05-07
and 1990-07-08. It is also known that the individual's 23rd birthday occurred at or before
2012-06-23, therefore their birth date window can be narrowed.

The `dobEstimator` package leverages this approach to accurately pinpoint the time interval
in which an individual's date of birth must fall.

The package works best when there are a large number of age observations per person, especially
when these are spread throughout the year and where there are > 1 distinct ages recorded.
If only a single observation is available, the date of birth interval will be one year wide.
It is only possible to identify a specific date of birth where two distinct ages are recorded
on consecutive calendar dates.

## Use cases

Once `dobEstimator` has estimated an individual's date of birth interval, it could be used for:

* Imputing missing age observations
* To improve the accuracy of age matching in case-control studies
* To estimate a distribution of dates of birth for a population

### Re-identification in anonymised datasets

This package __should not be used to attempt to re-identify individuals in anonymised datasets__.
Doing so is likely to breach privacy laws in many jurisdictions, and is likely to be against the terms of use of the dataset.
Users are strongly urged to adhere to all relevant privacy laws and ethical guidelines when utilising this package.

## Installation

You can install the development version of dobEstimator from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bairdj/dobEstimator")
```

## Example

Basic example of the [estimate_dob()] function using the above dataset:

```{r example}
library(dobEstimator)

print(age_dataset)

estimate_dob(age_dataset$age, age_dataset$date)
```

Note that [estimate_dob()] should only be used with data for a single individual.

If you need to calculate dates of births for multiple individuals, you should split
the age and date vectors by person first.

For example:


```{r multi-example-data, echo=FALSE}
multi_dataset <- data.frame(
  person_id = rep(c("A", "B"), each = 3),
  age = c(20, 21, 21, 24, 24, 26),
  date = as.Date(c("2010-01-04", "2010-05-03", "2010-12-25", "2012-04-05", "2012-08-02", "2014-07-03"))
)
```

```{r warning=FALSE}
print(multi_dataset)

library(dplyr, warn.conflicts = FALSE)
multi_dataset %>%
  group_by(person_id) %>%
  summarise(dob = estimate_dob(age, date))
```